\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-07 smurf_paper.tex
small typo
Remove the notes on additional background on performance etc
The paper is long and doesn't really need proof that it scales
with cores or details on how much memory is required. The Starlink
URLs should be enough to get the data and NDF is just a distraction.

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) predominantly
low-frequency noise sources that are correlated amongst detectors, and
to iterate this process along with estimates of the map. Significant
experimentation with different models for noise sources during \scuba\
commissioning required a highly flexible software framework, and a
configurable user interface. To achieve this goal, while minimizing
development time, it was decided to build SMURF as a Starlink package
\citep{2009ASPC..411..418J}, which provides access to a large suite of
libraries (including a commanding and messaging interface, astrometric
coordinate conversions and generation of standard WCS information,
file formats etc.). Furthermore, Starlink\footnote{\url{http://www.starlink.ac.uk}} is
open-source\footnote{\url{https://github.com/Starlink}},
and already used extensively at the Joint Astronomy Centre (host of the JCMT) for
many other systems \citep{jenness2011}, which helps with interoperability. Though
originally written in FORTRAN, many of the core Starlink libraries are
now ported to native C, or at least have a C interface. It was
therefore decided to develop SMURF in C as well (rather than C++, for
example, which would have required adding further dependencies to
Starlink), although we have taken an object-oriented approach. For
example, all of the data for a given noise model is encapsulated in a
C structure, and there is a standard interface for all functions that
handle models (member data and functions for the class,
respectively). In this way it is easy to extend SMURF with new
models. Parallelisation is incorporated in the most time-consuming
low-level routines using threads (e.g., performing Fast Fourier
Transforms, re-gridding the data), usually handling either independent
blocks of bolometers, or blocks of time, in each thread, depending on
the nature of the calculation. \textbf{... something about scalability
...}

\item Discuss SMURF configurability

\item Some sort of proof/demonstration showing why the algorithm works

\item Discuss performance (execution time, memory, disk usage etc.)

\item Where do you get SMURF/starlink, usage of NDF

\item Talk about pipeline?

@@

It was immediately found in the analysis of \scuba\ data that the
$\chi^2_\mathrm{r}$ test described above i not a sufficient stopping
criterion. The reason for this is that various components of the model
are, under normal circumstances, highly degenerate. One simple example
is the degeneracy between large-scale astronomical structures (larger
than the array footprint), and the common-mode rejection step
(\model{COM}), which will be illustrated in
Section~\ref{sec:point}. Since the flatness of the residual only
reflects the ability of the model to fit the shape of the bolometer
time-series, it does not necessarily correlate with convergence of the
most important model component: the map. Indeed, examination of the
map after large numbers of iterations (e.g., $>10$), shows the
presence of large-scale structures that are anti-correlated with
structures in other model components. For this reason, a second
map-based convergence test was added.

>>>>

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) predominantly
low-frequency noise sources that are correlated amongst detectors, and
to iterate this process along with estimates of the map. Significant
experimentation with different models for noise sources during \scuba\
commissioning required a highly flexible software framework, and a
configurable user interface. To achieve this goal, while minimizing
development time, it was decided to build SMURF as a Starlink package
\citep{2009ASPC..411..418J}, which provides access to a large suite of
libraries (including a commanding and messaging interface, astrometric
coordinate conversions and generation of standard WCS information,
file formats etc.). Furthermore, Starlink\footnote{\url{http://www.starlink.ac.uk}} is
open-source\footnote{\url{https://github.com/Starlink}},
and already used extensively at the Joint Astronomy Centre (host of the JCMT) for
many other systems \citep{jenness2011}, which helps with interoperability. Though
originally written in FORTRAN, many of the core Starlink libraries are
now ported to native C, or at least have a C interface. It was
therefore decided to develop SMURF in C as well (rather than C++, for
example, which would have required adding further dependencies to
Starlink), although we have taken an object-oriented approach. For
example, all of the data for a given noise model is encapsulated in a
C structure, and there is a standard interface for all functions that
handle models (member data and functions for the class,
respectively). In this way it is easy to extend SMURF with new
models. Parallelisation is incorporated in the most time-consuming
low-level routines using threads (e.g., performing Fast Fourier
Transforms, re-gridding the data), usually handling either independent
blocks of bolometers, or blocks of time, in each thread, depending on
the nature of the calculation.

@@

It was immediately found in the analysis of \scuba\ data that the
$\chi^2_\mathrm{r}$ test described above is not a sufficient stopping
criterion. The reason for this is that various components of the model
are, under normal circumstances, highly degenerate. One simple example
is the degeneracy between large-scale astronomical structures (larger
than the array footprint), and the common-mode rejection step
(\model{COM}), which will be illustrated in
Section~\ref{sec:point}. Since the flatness of the residual only
reflects the ability of the model to fit the shape of the bolometer
time-series, it does not necessarily correlate with convergence of the
most important model component: the map. Indeed, examination of the
map after large numbers of iterations (e.g., $>10$), shows the
presence of large-scale structures that are anti-correlated with
structures in other model components. For this reason, a second
map-based convergence test was added.
